---
title: "Assignment 2, Titanic Analysis"
author: "Robby Jeffries"
date: "9/25/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

```


### What actually happened on that fateful day?

On April 15th, 1912, the "unsinkable" happened. The RMS Titanic sank. Although detailed information exists for many of the passengers, many people have never explored the data surrounding this voyage. Let's take a closer look at this data and uncover the details of this historic event.

I referenced [this page](https://medium.com/analytics-vidhya/a-beginners-guide-to-learning-r-with-the-titanic-dataset-a630bc5495a8) as I considered my analysis.

```{r}
# STEP 1: IMPORT AND CLEAN THE DATA

## If necessary, install the following packages on your machine:
# install.packages("titanic")
# install.packages("tidyverse")
# install.packages("ggcorrplot")
# install.packages("dummies")
# install.packages("caret")
# install.packages("e1071")
# install.packages("ROCR")


# import libraries
library(titanic)
library(tidyverse)
library(ggcorrplot)
library(dummies)
library(caret)
library(e1071)
library(ROCR)

# import raw data
raw_data <- bind_rows(as_tibble(titanic::titanic_train), as_tibble(titanic::titanic_test))

# save the raw data to a data frame variable
clean_raw_data <- raw_data

# convert 'Pclass' to a categorical variable
clean_raw_data$Pclass = factor(clean_raw_data$Pclass, order=TRUE, levels = c(3, 2, 1))

# convert 'Survived' to a categorical variable, and change from "0" and "1" to a more readable "No" and "Yes"
# convert 'Sex' to a categorical variable, and capitalize "male" and "female"
clean_raw_data <- clean_raw_data %>% 
  mutate(Survived = factor(Survived)) %>%
  mutate(Survived = forcats::fct_recode(Survived, "No" = "0", "Yes" = "1")) %>%
  mutate(Sex = factor(Sex)) %>%
  mutate(Sex = forcats::fct_recode(Sex, "Male" = "male", "Female" = "female"))


######################################################
# Different variables contain different amounts  
# of na values. To account for this, I created a 
# new data frame for each different plot. 
#
# This isn't very efficient, and it's repetitive.
#
# After completing the assignment, I realized that
# I could have run the following command to see 
# how many na's there were in each variable:
#
# clean_raw_data %>% purrr::map_dbl(~sum(is.na(.)))
# 
# If I were to start from scratch, I would have only 
# created new data frames where necessary to maximize
# the script's computational efficiency. :)
######################################################


# create a data frame to plot the survival rate in general. Omit na values.
df1 <- clean_raw_data
df1 = select(df1, Survived)
df1 = na.omit(df1)

# total how many rows are in the raw data and how many rows are in the cleaned data
raw_count <- count(raw_data)
cleaned_count <- count(df1)

```
***
### Question 1: Who lived and who died, demographically?

Let's first take a look at how many people lived and how many people died. The data set we are analyzing contains information for `r raw_count` passengers. Of those `r raw_count`, the data set contains survival information for `r cleaned_count` passengers.

```{r}

ggplot(df1, aes(x = Survived)) +
  geom_bar(width=0.5, fill = "steelblue") +
  geom_text(stat='count', aes(label=stat(count)), vjust=-0.5) +
  ggtitle("Survival in General") +
  theme_minimal()

number_died <- sum(df1$Survived == "No")
number_survived <- sum(df1$Survived == "Yes")
```

***
### Question 2: Was gender a factor?

From the previous graph, we can see that **`r number_died`** people died and **`r number_survived`** people survived. Let's double-click on this stat and analyze the survival rate by sex.

```{r}

# create a new data frame to analyze survival rate by sex.
# this will ensure that rows of data are not dropped because 
# an na exists elsewhere in the row

df_sex <- clean_raw_data
  # select the columns that we are interested in analyzing
  df_sex = select(df_sex, Survived, Sex)
  # omit NA values
  df_sex = na.omit(df_sex)

ggplot(df_sex, aes(Sex, fill=Survived)) +
  geom_bar(position = "fill") + 
  ylab("Survival Rate") + 
  geom_hline(yintercept = (sum(df_sex$Survived=="Yes")/nrow(df_sex)), col = "white", lty = 2) + 
  ggtitle("Survival Rate by Sex") +
  scale_fill_manual(values = c("grey18", "springgreen3")) +
  theme_minimal()
```

As the above chart shows, the majority of women survived, and the majority of men died. The white horizontal line represents the average survival rate without regard to sex. This shows us that the men's survival rate was below-average, and the women's survival rate was above-average. This is consistent with our expectation because women and children were given priority on the lifeboats.

***
### Question 3: Was class of service a factor?

The passengers were in three different classes: 1st, 2nd, and 3rd. How many people died in each class?

```{r}
# create a new data frame to analyze survival rate by class
df_class <- clean_raw_data
  # select the columns that we are interested in analyzing
  df_class = select(df_class, Survived, Pclass)
  # omit NA values
  df_class = na.omit(df_class)
  
df_class <- df_class %>% 
  mutate(Pclass = forcats::fct_recode(Pclass, "3rd Class" = "3", "2nd Class" = "2", "1st Class" = "1")) 

ggplot(df_class, aes(x = Survived, fill=Pclass)) +
 geom_bar(position = position_dodge()) +
 geom_text(stat='count', 
           aes(label=stat(count)), 
           position = position_dodge(width=1), 
           vjust=-0.5) + 
  ggtitle("Survival by Passenger Class") +
  ylab("Count") +
  scale_fill_manual(values = c("#7F735F", "azure3", "#E9AF4E")) +
  theme_minimal()
```

As one might expect, the majority of passengers who died were in the lowest (3rd) class, and the majority of survivors were first-class passengers.



***
### Question 4: What other patterns do you see in the data?

#### What did the survival rate look like by age?

```{r}
# create a new data frame to analyze survival rate by age
df_age <- clean_raw_data
  # select the columns that we are interested in analyzing
  df_age = select(df_age, Survived, Age)
  # omit NA values
  df_age = na.omit(df_age)

# Discretize age to plot survival
df_age$Discretized.age = cut(df_age$Age, c(0,10,20,30,40,50,60,70,80,100))

# Plot discretized age
ggplot(df_age, aes(x = Discretized.age, fill=Survived)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat='count', 
            aes(label=stat(count)), 
            position = position_dodge(width=1), 
            vjust=-0.5) +
  ggtitle("Survival by Age") +
  xlab("Discretized Age") +
  ylab("Count") +
  scale_fill_manual(values = c("grey18", "springgreen3")) +
  theme_minimal()

# discard the temporary variable after using it.
df_age$Discretized.age = NULL

ggplot(df_age, aes(x = Age, fill = Survived)) +
  geom_density(alpha = 0.4) +
  ggtitle("Density Plot of Age related to Survival") +
  scale_fill_manual(values = c("grey18", "springgreen3"))


```

The density plot above shows whether survival or death was more common at a given age. That is, where the green ("Yes") color has the greater density, survival was more common. The survival rate was best among young children. Where the grey ("No") color has the greater density, death was more common. Passengers in their 20s were the most at-risk, and we can also see that the elderly also had a greater risk of dying.

***
### Question 5: What is something surprising in the data that others may have missed?

#### What did the survival rate look like by fare price?

When people think about the victims of the Titanic's sinking, they may intuitively associate age or status with survival. Let's add a level of granularity. Since this data set contains the price of each passenger's fare, let's check whether or not the amount of money that someone spent on their ticket impacted their survival.

Perhaps those that spent more money on their ticket were more cognizant of how to navigate the ship and claim a spot on a lifeboat? Let's take a look.

*Note: Although some passengers received a free ticket, most passengers spent money on their fare.*

```{r}
# create a new data frame to analyze survival rate by fare
df_fare <- clean_raw_data
  # select the columns that we are interested in analyzing
  df_fare = select(df_fare, Survived, Fare)
  # omit NA values
  df_fare = na.omit(df_fare)

# Discretize fare to plot survival
df_fare$Discretized.fare = cut(df_fare$Fare, c(-1,50,100,150,200,250,300,350,400,450,500,550,600,650))

# Plot discretized fare
ggplot(df_fare, aes(x = Discretized.fare, fill=Survived)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat='count', 
            aes(label=stat(count)), 
            position = position_dodge(width=1), 
            vjust=-0.5) +
  ggtitle("Survival by Fare") +
  xlab("Discretized Fare ($)") +
  ylab("Count") +
  scale_fill_manual(values = c("grey18", "springgreen3")) +
  theme_minimal()

# discard the temporary variable after using it.
df_fare$Discretized.fare = NULL

```

Interestingly, approximately two-thirds of the passengers who spent the least amount of money died in the chaos. As the fare price increased, we see that the majority of passengers survived. Further, all three passengers that spent between $500-$550 survived. The plot above contains a strong right-skew. To account for this, let's create a density plot on a log scale.

```{r}
ggplot(df_fare, aes(x = log(Fare), fill = Survived)) +
  geom_density(alpha = 0.4) + 
  xlab("log(Fare)") +
  ylab("Density") + 
  ggtitle("Density Plot of Fare related to Survival") +
  scale_fill_manual(values = c("grey18", "springgreen3")) 

```

Sure enough, the log scale shows us that the passengers with cheaper tickets had a significantly greater risk of dying. Further analysis would be required before causal inferences could be made, but a correlation between survival and fare certainly exists.


***
### Logistic Regression to predict survival

*Caveat: I did not know how to approach regression analysis off-hand; so I went to Google. I discovered [this page](https://rpubs.com/abhaypadda/logistic-regression-using-titanic-data), written by Abhay Padda. For practice, I replicated his code here. While much of Padda's code worked as written, several sections requried modifications to work with our data set.*

Logistic regressions are useful when the dependent variable is binary. In this case, our dependent variable is survival (0 = died, 1 = survived). When logistic functions are plotted, they resemble an elongated letter 's'. For a given value along the x-axis, the plot will tell you the probability that the passenger survived.

A question arises: at what threshold should a probability be classified as a survival or a death? 50%? 65%? That is where the [**Receiver Operating Characteristic (ROC) curve**](https://en.wikipedia.org/wiki/Receiver_operating_characteristic) comes into play. The ROC curve is built by (1) running a logistic regression and (2) computing a [confusion matrix](https://en.wikipedia.org/wiki/Confusion_matrix) for each individual passenger, which defines a new threshold of survival for each passenger.

The end goal is to strike an optimal balance between the true positive rate and the false positive rate. The following graph shows the ROC curve for the Titanic data set, and the optimal configuration is near the top left corner. As we might have intuitively guessed for a simple data set such as this, a good probability threshold for predicting survival is approximately 50%. 

*(Note: this is not predicting that 50% of passengers survived. It simply says that for a logistic regression of survival, passengers with a probability of survival >=50% likely survived and passengers with a probability of survival <=50% likely died.)*

```{r}

# load the datasets
data("titanic_train")
data("titanic_test")

# setting survived column for test data to NA
titanic_test$Survived <- NA

# combining training and testing datasets
complete_data <- rbind(titanic_train, titanic_test)

# missing values imputation
complete_data$Embarked[complete_data$Embarked==""] <- "S"

complete_data$Age[is.na(complete_data$Age)] <- median(complete_data$Age, na.rm=T)

# Removing cabin as it has very high missing values, passengerID, Ticket, and Name are not required
titanic_data <- complete_data %>% dplyr::select(-c(Cabin, PassengerId, Ticket, Name))

# converting "Survived", "Pclass", "Sex", "Embarked" to factors
for (i in c("Survived", "Pclass", "Sex", "Embarked")){ 
  titanic_data[,i] = as.factor(titanic_data[,i])
  }

# create dummy variables for categorical variables
titanic_data <- dummies::dummy.data.frame(titanic_data, names = c("Pclass", "Sex", "Embarked"), sep = "_")

# Splitting training and test data
train <- titanic_data[1:667,]
test <- titanic_data[668:889,]

# Model creation
model <- stats::glm(Survived ~., family = binomial(link = 'logit'), data = train)

# Model Summary
# summary(model)

## Using anova() to analyze the table of deviance
# anova(model, test = "Chisq")

# Predicting test data
result <- predict(model, newdata = test, type = 'response')
result <- ifelse(result > 0.5, 1, 0)

## Confusion matrix and statistics, which 
#caret::confusionMatrix(data = as.factor(result), reference = as.factor(test$Survived))

# Receiver Operating Characteristic (ROC) Curve and calculating the area under the curve (AUC)
predictions <- predict(model, newdata = test, type = "response")

ROCRpred <- ROCR::prediction(predictions, test$Survived)

ROCRperf <- ROCR::performance(ROCRpred, measure = "tpr", x.measure = "fpr")

plot(ROCRperf, colorize = TRUE, text.adj = c(-0.2, 1.7), print.cutoffs.at = seq(0, 1, .1), main="Receiver Operating Characteristic Curve")

auc <- performance(ROCRpred, measure = "auc")

auc <- round(auc@y.values[[1]], digits = 2)

auc_perc <- auc * 100

```

The area under the curve (AUC) is `r auc`. This value is meaningful if you want to compare multiple ROC curves and determine which one is the best model. The greater the AUC, the better the model. An area of 1.00 or 100% is a perfect model. Therefore, I conclude that the AUC of the above graph is 'pretty good' with an area of `r auc_perc`%.

For additional information regarding ROC and AUC, I watched this [StatQuest with Josh Starmer](https://www.youtube.com/watch?v=4jRBRDbJemM).